---
# PUBPOL599B Final Project
## Emma Weaver
### March 6, 2018
---

Get dataset with region data from GitHub:
```{r,eval=TRUE}
linkRegions= 'https://github.com/efweavs1989/PUBPOL599B_DATA/blob/master/hh_sec_a.csv?raw=true'

region=read.csv(linkRegions)
```

Get dataset with Section I data from Github:
```{r,eval=TRUE}
i_link= 'https://github.com/efweavs1989/PUBPOL599B_DATA/raw/master/hh_sec_i.csv'
i_data= read.csv(i_link)
```

Get dataset with Section Q1 data from Github:
```{r,eval=TRUE}
q1_link= 'https://github.com/efweavs1989/PUBPOL599B_DATA/raw/master/hh_sec_q1.csv'
q1_data= read.csv(q1_link)
```

View names to identify which variables need to be kept. Keep household ID, region number, and region name:
```{r,eval=TRUE}
names(region)
```
```{r,eval=TRUE}
region=region[,c("y4_hhid","hh_a01_1", "hh_a01_2")]
```

View names for Section I data to determine which to keep:
```{r,eval=TRUE}
names(i_data)
```

Select variables from Section I to anlayze. These are:

"hh_i05": In the past year, how much have you paid on repairs to your dwelling? (TSH)
"hh_i06": In the past year, how much have you paid in improvements to your home (excluding any purchases listed in previous question)? (TSH)
"hh_i08": The walls of the main dwelling are predominantly made of what materials? (See questionnaire for options)
"hh_i09": The roof of the main dwelling is predominantly made of what materials? (See questionnaire for options)
"hh_i10": The floor of the main dwelling is predominantly made of what materials? (See questionnaire for options)
```{r,eval=TRUE}
i_data=i_data[,c("y4_hhid","hh_i05","hh_i06","hh_i08","hh_i09","hh_i10")]
```

View names from Section Q1 data to determine which namesto keep:
```{r,eval=TRUE}
names(q1_data)
```

Select variables from Section Q that want to anlayze. These are:

"hh_q01_1": Did you or anyone in your household use M-PESA to transfer money over the last 12 months
"hh_q01_2": Did you or anyone in your household use EZY PESA to transfer money over the last 12 months
"hh_q01_3": Did you or anyone in your household use Airtel Money to transfer money over the last 12 months
"hh_q01_4": Did you or anyone in your household use TIGO PESA to transfer money over the last 12 months

```{r,eval=TRUE}
q1_data=q1_data[,c("y4_hhid","hh_q01_1","hh_q01_2","hh_q01_3","hh_q01_4")]
```

Merge data sets. Need to merge one by one. 

```{r,eval=TRUE}
i_reg=merge(region,i_data)
projectData=merge(i_reg,q1_data)
```

Check the data type to determine how to aggregate the data.

```{r,eval=TRUE}
str(projectData)
```


Rename the variables of interest:
```{r,eval=TRUE}
names(projectData)=gsub("y4_hhid", "Household_ID", names(projectData))
names(projectData)=gsub("hh_a01_1", "Region_Code", names(projectData))
names(projectData)=gsub("hh_a01_2", "Region_Name", names(projectData))
names(projectData)=gsub("hh_i05", "Value_of_Repairs", names(projectData))
names(projectData)=gsub("hh_i06", "Value_of_Improvements", names(projectData))
names(projectData)=gsub("hh_i08", "Wall_Material", names(projectData))
names(projectData)=gsub("hh_i09", "Roof_Material", names(projectData))
names(projectData)=gsub("hh_i10", "Floor_Material", names(projectData))
names(projectData)=gsub("hh_q01_1", "MPesa_Use", names(projectData))
names(projectData)=gsub("hh_q01_2", "Ezy_Pesa_Use", names(projectData))
names(projectData)=gsub("hh_q01_3", "Airtel_Use", names(projectData))
names(projectData)=gsub("hh_q01_4", "Tigo_Pesa_Use", names(projectData))
```

Double check the names in projectData:
```{r}
names(projectData)
```

The primary variable of interest is the value of improvements (how much each household paid for improvements to their home in the past year). 

First, aggregate all variables that can be aggregated by the mean:
```{r,eval=TRUE}
(mean_variables=aggregate(cbind(Value_of_Repairs, Value_of_Improvements, MPesa_Use, Ezy_Pesa_Use, Airtel_Use, Tigo_Pesa_Use) ~ Region_Code, data=projectData, FUN=mean))
```
The mobile money use is measured as either 1 = do not use this service, or 2 = use this service.

Now aggregate those variables that need to be aggregated by mode. To do this we need to first create a function which calculates the mode of each variable: 
```{r,eval=TRUE}
getMode=function(aColumn){
  freqTable=table(aColumn)
  maxFrequency=max(freqTable)
  names(freqTable[freqTable==maxFrequency])[1]
}
```

Now that we have created this function use it to aggregate all categorical variables:
```{r,eval=TRUE}
(mode_variables=aggregate(cbind(Wall_Material, Roof_Material, Floor_Material) ~ Region_Code, data=projectData, FUN=getMode))
```

Now merge these two aggregated files together:
```{r,eval=TRUE}
projectData=merge(mean_variables,mode_variables)
```

The region code currently skips from 25 to 51. Revise this numbering below:
```{r}
projectData$Region_Code=gsub(51, 26, projectData$Region_Code)
projectData$Region_Code=gsub(52, 27, projectData$Region_Code)
projectData$Region_Code=gsub(53, 28, projectData$Region_Code)
projectData$Region_Code=gsub(54, 29, projectData$Region_Code)
projectData$Region_Code=gsub(55, 30, projectData$Region_Code)
projectData$Region_Code
```

And change this variable back to numeric:
```{r}
projectData$Region_Code=as.numeric(projectData$Region_Code)
projectData$Region_Code
```

Returning to our primary variable of interest, we will begin to perform univariate analysis to further explore this variable. 

First, look at the summary statistics:
```{r}
summary(projectData$Value_of_Improvements)
```

Now, find the standard deviation:
```{r,eval=TRUE}
sd(projectData$Value_of_Improvements,na.rm = T)

```

Skewness:
```{r,eval=TRUE}
library(moments)
skewness(projectData$Value_of_Improvements,na.rm = T)
```

Kurtosis:
```{r,eval=TRUE}
kurtosis(projectData$Value_of_Improvements,na.rm = T)
```

Now create a histogram:
```{r,eval=TRUE}
library(ggplot2)
# keeping non missing:
data=projectData[is.finite(projectData$Value_of_Improvements),]
#
# selecting a var:
var=data$Value_of_Improvements
#
# saving mean and sd:
mnVar=mean(var,na.rm = T)
sdVar=sd(var,na.rm = T)
#
#plotting
base = ggplot(data, aes(x=var))
hist = base + geom_histogram(fill="green", 
                             color='grey',
                          aes(y=..density..))
#
histAndNormal = hist + stat_function(fun=dnorm,
                                     color="red",
                                     args=list(mean=mnVar,sd=sdVar))
#
histAndNormal
```

Bivariate analysis:
I will explore the relationship between the amount spent on improvements and secondary, numerical variables, by region.

```{r}
str(projectData)
```

Compute the correlation between the value of improvments and other numerical variables:
```{r,eval=TRUE}
res <- cor(projectData[,c(2:7)],use='complete.obs')
round(res, 2)
```

Below is a visual representation of the correlation between value of improvments, value of repairs and use of one mobile money platform (MPesa):
```{r,eval=TRUE}
library(car)
scatterplotMatrix(projectData[,c(2:4)], 
   main="Simple Scatterplot Matrix")
```

Next we should explore the relationship between some of our categorical variables and our main numeric variable of interest, the value of improvments.

Let's start by looking at at the relationship between the value of improvements and roof material. First, create a new variable which includes the names of the roofing material. 

```{r}
projectData$Roof_Material_Name[projectData$Roof_Material==1] <- "Grass, Leaves, Bamboo"
projectData$Roof_Material_Name[projectData$Roof_Material==4] <- "Metal Sheets"
```

Next, create a box plot:
```{r}
projectData=projectData[complete.cases(projectData),] 
p <- ggplot(projectData, aes(Roof_Material_Name,Value_of_Improvements))

p + geom_boxplot() + labs(title="The Value of Improvements by Roof Material") + scale_x_discrete(name = "Roof Material") +
        scale_y_continuous(name = "Value of Improvements\nMade to the Home")
```

Now create a new variable with the names of the wall material:
```{r}
projectData$Wall_Material_Name[projectData$Wall_Material==2] <- "Poles & Mud/Mud & Stones"
projectData$Wall_Material_Name[projectData$Wall_Material==4] <- "Mud Bricks"
projectData$Wall_Material_Name[projectData$Wall_Material==5] <- "Baked/Burnt Bricks"
projectData$Wall_Material_Name[projectData$Wall_Material==6] <- "Concrete, Cement, Stones"
```

And create a box plot:
```{r}
projectData=projectData[complete.cases(projectData),] 
p <- ggplot(projectData, aes(Wall_Material_Name,Value_of_Improvements))

p + geom_boxplot() + labs(title="The Value of Improvements by Wall Material")+scale_x_discrete(name = "Wall Material") +
        scale_y_continuous(name = "Value of Improvements\nMade to the Home")
```

Lastly, create a new variable with the names of the floor material:
```{r}
projectData$Floor_Material_Name[projectData$Floor_Material==1] <- "Earth"
projectData$Floor_Material_Name[projectData$Floor_Material==2] <- "Concrete, Cement, Tiles, Timber"
```

And create a box plot:
```{r}
projectData=projectData[complete.cases(projectData),] 
p <- ggplot(projectData, aes(Floor_Material_Name,Value_of_Improvements))

p + geom_boxplot() + labs(title="The Value of Improvements by Floor Material")+scale_x_discrete(name = "Floor Material") +
        scale_y_continuous(name = "Value of Improvements\nMade to the Home")
```

Clustering:
We will cluster variables by region and first need to confirm that every region only appears once.
```{r}
length(projectData$Region_Code)==length(unique(projectData$Region_Code))
```
Now make the region the row name:
Note for Emma: What happened to Region Name? Where did it get dropped? Where do I need to add it back?
```{r}
row.names(projectData)=projectData$Region_Code
```

Now double check the data of interest to make sure it is all the correct type:
```{r}
str(projectData[,c(1:7)])
```

We may be concerned that use of different mobile money platforms are highly correlated, which would impact our clustering. 
```{r}
res2 <- cor(projectData[,c(4:7)],use='complete.obs')
round(res2, 2)
```
The above shows us that Airtel use and Tigo Pesa use are highly correlated, and that MPesa use and Ezy Pesa use are highly correlated. Because of this we will conduct clustering with only MPesa and Airtel use variables. 

First we will produce a similarity matrix: 
```{r}
library(cluster)

demoSimi=daisy(projectData[,c(1:4,6)],metric = "euclidean",
               stand=TRUE) # standardizing
```

apply the clustering technique:
```{r}
demoClus=hclust(demoSimi,method = 'average') 
```

Now compare visually:
```{r}
plot(demoClus,cex=0.5)
```

Let's try to cluster in 5 groups:
```{r}
plot(demoClus,cex=0.3)
rect.hclust(demoClus, k = 5,border = c('orange','blue','red','green','purple'))
```

Now let's compare numerically:
```{r}
demsil5=silhouette(cutree(demoClus, k = 5), demoSimi)
clust5solution=data.frame(matrix(demsil5,ncol = 3,dimnames =dimnames(demsil5)))
row.names(clust5solution)=projectData$Region_Code
clust5solution
```

Let's incorporate the cluster into our original data:
```{r}
# saving solution chosen
solution=clust5solution

# creating country column
solution$Region_Code=row.names(clust5solution)
row.names(solution)=NULL # resetting rownames

# merging
projectData=merge(projectData,solution)
```

Now let's check the updated data frame:
```{r}
str(projectData)
```

Let's look at some of our variables of interest by cluster:
```{r}
palette=c('blue','orange','red','green','purple')
scatterplotMatrix(projectData[,c(2:7)], diagonal='histogram',reg.line=NULL , smoother=NULL,groups = projectData$cluster,col = palette,legend.plot=F)
```

Mapping Data:
To map this data, we first need to bring in the shapefile for the regions:

```{r}
compressedMap= "https://github.com/efweavs1989/PUBPOL599B_Project/raw/master/TZA_adm_shp.zip"
```

```{r}
library(utils)
temp=tempfile()
download.file(compressedMap, temp)
unzip(temp)
```
Below we check which shapefiles are currently in our directories:
```{r}
(maps=list.files(pattern = 'shp'))
```

We are interested in "TZA_adm1" because this is for the country and region level only.

Now we select that map:
```{r}
library(rgdal)
TZMap <- rgdal::readOGR("TZA_adm1.shp",stringsAsFactors=F) # use name
```

```{r}
names(TZMap)
```

Just keep the region code data:
```{r}
TZMap=TZMap[,c('ID_1')]
```

Let's check the data type of the column which includes the region code:
```{r}
str(TZMap$ID_1)
```

This is an integer in our other data, so let's convert this to an integer:
```{r}
TZMap$ID_1=as.numeric(TZMap$ID_1)
```

```{r}
str(TZMap$ID_1)
```

Now let's merge:
```{r}
contribTZmap=merge(TZMap,projectData, by.x='ID_1', by.y='Region_Code',all.x=F)
```

Install the packages required for managing color and divisions when plotting:
```{r}
library(RColorBrewer)
library(classInt)
```

First, let's plot the clusters created previously:
```{r}
varToPLot=projectData$cluster
```

Get colors and intervals
```{r}
numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)
```

Plot:
```{r}
legendText="clusters"
shrinkLegend=0.4
title="Plot of Variable Clusters"

plot(contribTZmap,col='purple',main=title)
plot(contribTZmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

Now plot the value of improvements:
```{r}
varToPLot=projectData$Value_of_Improvements
```

Get colors and intervals
```{r}
numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)
```
Plot:
```{r}
legendText="value of improvements"
shrinkLegend=0.4
title="Amount HH Spent on Improvements to the Home"

plot(contribTZmap,col='purple',main=title)
plot(contribTZmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

Now plot the value of repairs:
```{r}
varToPLot=projectData$Value_of_Repairs
```

Get colors and intervals
```{r}
numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)
```
Plot:
```{r}
legendText="value of repairs"
shrinkLegend=0.4
title="Amount HH Spent on Repairs to the Home"

plot(contribTZmap,col='purple',main=title)
plot(contribTZmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```
Plot the use of MPesa, one of the mobile money platforms used in the survey:
```{r}
varToPLot=projectData$MPesa_Use
```

Get colors and intervals
```{r}
numberOfClasses = 5
colorForScale='YlGnBu'
colors = brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, 
                            style = "quantile",
                            dataPrecision=2)
colorPallette <- findColours(intervals, colors)
```
Plot:
```{r}
legendText="MPesa Use"
shrinkLegend=0.4
title="Does the HH Use Mobile Money (MPesa)"

plot(contribTZmap,col='purple',main=title)
plot(contribTZmap, col = colorPallette,border='grey',add=T)

legend('topright', legend = names(attr(colorPallette, "table")), 
       fill = attr(colorPallette, "palette"), cex = shrinkLegend, 
       bty = "n",
       title=legendText)
```

